Do fast stellar centroiding methods saturate the Cramer-Rao lower bound?

### 2016-09-23 ms/ms.tex
edited the title and abstract

\title{Fast and optimal centroiding of faint stars}

\begin{abstract}
One of the most demanding tasks in astronomical image processing---in terms of precision---is the centroiding of stars. 
Upcoming large surveys are going to take images of billions of point sources, 
including many faint stars, with short exposure times. Real-time estimation of the
centroids of stars is crucial for real-time PSF estimation, and maximal precision is required
for measurements of proper motion. In this work, we aim to compare the performance of
various centroiding methods, when they are applied
to relatively low signal-to-noise ratio unsaturated stars.
In order to investigate how information-preserving these techniques are, we compare
the root-mean-squared-error with the fundamental Cram\'{e}r-Rao
lower bound assuming zero-mean constant Gaussian noise. We discuss two general circumstances in centroiding of faint stars: (i) when we have a good estimate
of the PSF, (ii) when we do not know the PSF. In the case that we know the PSF, 
we show that a fast polynomial centroiding after smoothing the image by the PSF can be 
as accurate as full PSF profile fitting. In the case that we do not
know the PSF, we demonstrate that although polynomial centroiding is not as
accurate as PSF profile fitting, it comes close to saturating the Cram\'{e}r-Rao lower bound
in a wide range of conditions. We also show that the moment-based method of center-of-light never comes close to 
saturating the bound, and thus it does not deliver reliable estimates of centroids.    
\end{abstract}

>>>>

\title{Do fast stellar centroiding methods saturate the Cramer-Rao lower bound?}

\begin{abstract}
One of the most demanding tasks in astronomical image processing---in terms of precision---is 
the centroiding of stars. Upcoming large surveys are going to take images of 
billions of point sources, including many faint stars, with short exposure times. 
Real-time estimation of the centroids of stars is crucial for real-time PSF estimation, 
and maximal precision is required for measurements of proper motion.

The fundamental Crame\'{e}r-Rao lower bound sets a limit on the root-mean-squared-error 
achievable by optimal estimators. Information-preserving estimators are able to deliver 
root-mean-squared-errors lower than the Crame\'{e}r-Rao bound. In this work, we aim to compare 
the performance of various centroiding methods, in terms of saturating the bound, when they 
are applied to relatively low signal-to-noise ratio unsaturated stars assuming zero-mean 
constant Gaussian noise. In order to make this comparison, we present the root-mean-squared-errors of 
these estimators and their corrsponding Crame\'{e}r-Rao bound as a function of the signal-to-noise ratio 
and the full-width at half-maximum of faint stars.

We discuss two general circumstances in centroiding of faint stars: (i) when we have a good estimate
of the PSF, (ii) when we do not know the PSF. In the case that we know the PSF, 
we show that a fast polynomial centroiding after smoothing the image by the PSF can be 
as information-reserving as the maximum-likelihood estimator with full PSF profile fitting. 
In the case that we do not know the PSF, we demonstrate that although polynomial centroiding 
is not as optimal as PSF profile fitting, it comes very close to saturating the Cram\'{e}r-Rao lower bound
in a wide range of conditions. We also show that the moment-based method of center-of-light 
never comes close to saturating the bound, and thus it does not deliver reliable estimates of centroids.    
\end{abstract}
